### 2024-03-16 README.Rmd
add CITATION.cff and udpate README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Whole-body connectome of a segmented annelid larva"
```

# About the project: 
- this is a repository of an R project containing code and source data for the Platynereis three-day-old larva connectome paper. Contributed by [Jékely lab](https://www.cos.uni-heidelberg.de/en/research-groups/gaspar-jekely).

This repository contains the data and code for our paper:

> Gáspár Jékely, Sanja Jasek, Martin Gühmann, Luis Alberto Bezares Calderón, Elizabeth A. Williams, Réza Shahidi (2024)
 \. _`r Title`_. Name of journal/book <>

Our pre-print is online here:

> Gáspár Jékely, Sanja Jasek, Martin Gühmann, Luis Alberto Bezares Calderón, Elizabeth A. Williams, Réza Shahidi (2024). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Gáspár Jékely, Sanja Jasek, Martin Gühmann, Luis Alberto Bezares Calderón, Elizabeth A. Williams, Réza Shahidi, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The project directory contains:

  - [:file\_folder: text](/Manuscript.qmd): Quarto source document
    for manuscript. Includes text, references and inserts the latest version of the figures from the /Figures folder. 
There is also a rendered html version,
    `Manuscript.html`, suitable for reading.

  - [:file\_folder: data](/data): Data used in the analysis.
  - [:file\_folder: scripts](/code): scripts used in the analysis.
  - [:file\_folder: figures](/pictures): Plots and other
    illustrations
  - [:file\_folder:
    supplements](/supplements):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.
  - [:file\_folder:
    source_data](/source_data):
    Source data files associated with each figure containing all data points that have been plotted in the figure.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

If you would like to use the template, you could open Rstudio in your computer, then go to File > New Project > Version Control > Git. Under repository URL add the URL of this page https://github.com/JekelyLab/Jekely_et_al_2024_connectome and save it in a local folder.

The R project will live in the folder (as working directory) where the Jekely_et_al_2024_connectome.Rproj file is saved. All other files and directories will be accessed relative to this working directory. This way the project is portable and self-contained.

The working directory contains this README file, a LICENCE file, a CITATION.cff file, the .gitignore file and the R project file. You can update these files to fit your project.

### Licenses

**Text, figures, code, data :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

### Acknowledgements

This project is modified after
https://annakrystalli.me/rrresearch/10_compendium.html