### 2024-09-25 supplements/Supplementary_file_cell_type_compendium.qmd
celltype compendium: new suppl website

---
title: "Compendium of all cell types in the three-day-old *Platynereis dumerilii* larva"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
---

```{r}
#| echo: false
#| include: FALSE
library(dplyr)
library(rio)
library(tinytable)
library(knitr)
library(gt)
library(catmaid)
library(rgl)
library(htmltools)

# catmaid connection
conn_http1 <- catmaid_login(
  server="https://catmaid.jekelylab.ex.ac.uk/", 
  authname="AnonymousUser",
  config=httr::config(ssl_verifypeer=0, http_version=1)
)
```

# Neuronal cell types

## Sensory neurons

```{r}
#| echo: false

plot_background <- function(x){
  nopen3d() # opens a pannable 3d window
  par3d(windowRect = c(0, 0, 200, 200)) #resize for frontal view
  plot3d(bounding_dots, add=T, col="white") 
  plot3d(yolk, add=T, alpha=0.1, col="#E2E2E2") 
  plot3d(acicula, soma=F, lwd=2, add=T, alpha=1, col="grey70")
  par3d(zoom=0.52)
  nview3d("frontal", extramat=rotationMatrix(0.2, 1, 0.1, 0.5))
  #z-axis clip
  #clipplanes3d(0, 0, -1, 65000)
  #y-axis clip
  clipplanes3d(1, 0, 0.16, 7000)
  #x-axis clip
  clipplanes3d(0, -1, 0.16, 130000)
}

plot_background_ventral <- function(x){
  nopen3d() # opens a pannable 3d window
  par3d(windowRect = c(0, 0, 200, 200)) #resize for frontal view
  plot3d(bounding_dots, add=T, col="white") 
  plot3d(yolk, add=T, alpha=0.1, col="#E2E2E2") 
  plot3d(acicula, soma=F, lwd=2, add=T, alpha=1, col="grey70")
  par3d(zoom=0.6)
  nview3d('ventral', extramat=(rotationMatrix(0.35, 1, 0, 0)%*%rotationMatrix(0.05, 0, 0, 1)))
  #y-axis clip
  clipplanes3d(1, 0, 0.16, 7000)
  #x-axis clip
  clipplanes3d(0, -1, 0.16, 130000)
}

celltypes <- rio::import("Supplementary_Table1.csv")

#load anatomy landmarks
bounding_dots = nlapply(read.neurons.catmaid("^bounding_dots$", pid=11),
                          function(x) smooth_neuron(x, sigma=6000))
yolk <- catmaid_get_volume(4, rval = c("mesh3d", "catmaidmesh", "raw"),
                             invertFaces = T, conn = NULL, pid = 11)
acicula <-   nlapply(read.neurons.catmaid("^acicula$", pid=11),
                       function(x) smooth_neuron(x, sigma=6000))

#functions for table and neuron plot
plot_cell_name <- function(annotation, type_of_cell){
  name <- celltypes |>
  filter(`Sensory/inter/motor neuron` == type_of_cell) |> 
  filter(`CATMAID annotation` == annotation) |> 
  pull(`CATMAID annotation`, `name of cell type`)

kable(name, col.names = c("name of cell type", "CATMAID annotation"))
}

plot_cell_chars <- function(annotation, type_of_cell){
  celltypes |>
  filter(`Sensory/inter/motor neuron` == type_of_cell) |>
  filter(`CATMAID annotation` == annotation) |>   
  select(`number of cells`, `soma position`, `region`, `transmitter phenotype`, `postsynaptic partners (<10 synapses)`, `presynaptic partners (<10 synapses)`, `reference`) |>
  kable()
}

#function to plot rgl widget with neuron
plot_cell_rgl <- function(annotation, type_of_cell, background){
  ID <- celltypes |>
    filter(`Sensory/inter/motor neuron` == type_of_cell) |> 
    filter(`CATMAID annotation` == annotation) |> 
    pull(`CATMAID annotation`)

  neuron1 = nlapply(read.neurons.catmaid(ID, pid=11),
                    function(x) smooth_neuron(x, sigma=6000))
  if(background == "ventral"){
    plot_background_ventral()
  }
  else{
    plot_background()
  }
  plot3d(neuron1, WithConnectors = F, soma=TRUE, lwd=3,
         add=T, alpha=0.7, col="#0072B2")

  s <- scene3d()
  rglwidget(s, width = 400, height = 100)
}

```

### Photoreceptors

#### Rhabdomeric photoreceptors
```{r}
#| echo: false
# PRC
annot <- "celltype1"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```
[3D visualisation: ](`r path`)

```{r}
#| echo: false
#| eval: false
# eyespot-PRCR3
annot <- "celltype33"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)
```

```{r}
#| echo: false
#| eval: false
# eyespot-PRCR1
annot <- "celltype34"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)
```

#### Ciliary photoreceptors
```{r}
#| echo: false
#| eval: false
# cPRC
annot <- "celltype5"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)
```

### Mechanoreceptors

#### CR (collar receptor) hydrodynamic vibration-receptor neurons

```{r}
#| echo: false
#| eval: false
# hCR
annot <- "celltype100"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)
```

```{r}
#| echo: false
#| eval: false
# ventralpygCR
annot <- "celltype101"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")

widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)
```

```{r}
#| echo: false
#| eval: false
# doCRunp
annot <- "celltype102"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)
```

#### PU (penetrating uniciliated) mechanosensory neurons

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype88")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype89")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype90")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype91")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype92")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype93")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype94")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype95")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype96")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype97")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

```{r}
#| echo: false
#| eval: false
# PU neurons celltype88-98
annot <- ("celltype98")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
close3d()

path <- paste("../../supplements/celltype_RGLs/", annot, ".html", sep = "")
htmltools::save_html(widgets, path)

```

### Chemoreceptors