### 2024-09-25 supplements/celltype_compendium_website/index.qmd

---
title: "*Platynereis dumerilii* three-day-old connectome --- cell-type compendium"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
---

```{r}
#| echo: false
#| include: FALSE
library(dplyr)
library(rio)
library(tinytable)
library(knitr)
library(gt)
library(catmaid)
library(rgl)
library(htmltools)

# catmaid connection
conn_http1 <- catmaid_login(
  server="https://catmaid.jekelylab.ex.ac.uk/", 
  authname="AnonymousUser",
  config=httr::config(ssl_verifypeer=0, http_version=1)
)
```

# Neuronal cell types

## Sensory neurons

```{r}
#| echo: false

plot_background <- function(x){
  nopen3d() # opens a pannable 3d window
  par3d(windowRect = c(0, 0, 200, 200)) #resize for frontal view
  plot3d(bounding_dots, add=T, col="white") 
  plot3d(yolk, add=T, alpha=0.1, col="#E2E2E2") 
  plot3d(acicula, soma=F, lwd=2, add=T, alpha=1, col="grey70")
  par3d(zoom=0.52)
  nview3d("frontal", extramat=rotationMatrix(0.2, 1, 0.1, 0.5))
  #z-axis clip
  #clipplanes3d(0, 0, -1, 65000)
  #y-axis clip
  clipplanes3d(1, 0, 0.16, 7000)
  #x-axis clip
  clipplanes3d(0, -1, 0.16, 130000)
}

plot_background_ventral <- function(x){
  nopen3d() # opens a pannable 3d window
  par3d(windowRect = c(0, 0, 200, 200)) #resize for frontal view
  plot3d(bounding_dots, add=T, col="white") 
  plot3d(yolk, add=T, alpha=0.1, col="#E2E2E2") 
  plot3d(acicula, soma=F, lwd=2, add=T, alpha=1, col="grey70")
  par3d(zoom=0.6)
  nview3d('ventral', extramat=(rotationMatrix(0.35, 1, 0, 0)%*%rotationMatrix(0.05, 0, 0, 1)))
  #y-axis clip
  clipplanes3d(1, 0, 0.16, 7000)
  #x-axis clip
  clipplanes3d(0, -1, 0.16, 130000)
}

celltypes <- rio::import("../Supplementary_Table1.csv")

#load anatomy landmarks
bounding_dots = nlapply(read.neurons.catmaid("^bounding_dots$", pid=11),
                          function(x) smooth_neuron(x, sigma=6000))
yolk <- catmaid_get_volume(4, rval = c("mesh3d", "catmaidmesh", "raw"),
                             invertFaces = T, conn = NULL, pid = 11)
acicula <-   nlapply(read.neurons.catmaid("^acicula$", pid=11),
                       function(x) smooth_neuron(x, sigma=6000))

#functions for table and neuron plot
plot_cell_name <- function(annotation, type_of_cell){
  name <- celltypes |>
  filter(`Sensory/inter/motor neuron` == type_of_cell) |> 
  filter(`CATMAID annotation` == annotation) |> 
  pull(`CATMAID annotation`, `name of cell type`)

kable(name, col.names = c("name of cell type", "CATMAID annotation"))
}

plot_cell_chars <- function(annotation, type_of_cell){
  celltypes |>
  filter(`Sensory/inter/motor neuron` == type_of_cell) |>
  filter(`CATMAID annotation` == annotation) |>   
  select(`number of cells`, `soma position`, `region`, `transmitter phenotype`, `postsynaptic partners (<10 synapses)`, `presynaptic partners (<10 synapses)`, `reference`) |>
  kable()
}

#function to plot rgl widget with neuron
plot_cell_rgl <- function(annotation, type_of_cell, background){
  ID <- celltypes |>
    filter(`Sensory/inter/motor neuron` == type_of_cell) |> 
    filter(`CATMAID annotation` == annotation) |> 
    pull(`CATMAID annotation`)

  neuron1 = nlapply(read.neurons.catmaid(ID, pid=11),
                    function(x) smooth_neuron(x, sigma=6000))
  if(background == "ventral"){
    plot_background_ventral()
  }
  else{
    plot_background()
  }
  plot3d(neuron1, WithConnectors = F, soma=TRUE, lwd=3,
         add=T, alpha=0.7, col="#0072B2")

  s <- scene3d()
  rglwidget(s, width = 400, height = 100)
}

```

### Photoreceptors

#### Rhabdomeric photoreceptors

```{r}
#| echo: false
# PRC
annot <- "celltype1"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```
![](snapshots/celltype1.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype1.html){target="_blank"}

```{r}
#| echo: false

# eyespot-PRCR3
annot <- "celltype33"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype33.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype33.html){target="_blank"}

```{r}
#| echo: false

# eyespot-PRCR1
annot <- "celltype34"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype34.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype34.html){target="_blank"}

#### Ciliary photoreceptors

```{r}
#| echo: false

# cPRC
annot <- "celltype5"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype5.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype5.html){target="_blank"}

### Mechanoreceptors

#### CR (collar receptor) hydrodynamic vibration-receptor neurons

```{r}
#| echo: false

# hCR
annot <- "celltype100"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype100.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype100.html){target="_blank"}

```{r}
#| echo: false

# ventralpygCR
annot <- "celltype101"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")

widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype101.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype101.html){target="_blank"}

```{r}
#| echo: false

# doCRunp
annot <- "celltype102"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype102.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype102.html){target="_blank"}

#### PU (penetrating uniciliated) mechanosensory neurons

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype88")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype88.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype88.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype89")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype89.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype89.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype90")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype90.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype90.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype91")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype91.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype91.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype92")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype92.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype92.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype93")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype93.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype93.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype94")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype94.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype94.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype95")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype95.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype95.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype96")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype96.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype96.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype97")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype97.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype97.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype98")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype98.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype98.html){target="_blank"}

#### PB (penetrating biciliated) mechanosensory neurons

```{r}
#| echo: false

# PB neurons celltype80 and 54
annot <- ("celltype80")
plot_cell_name(annot, "Sensory neuron Interneuron")
plot_cell_chars(annot, "Sensory neuron Interneuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron Interneuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype80.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype80.html){target="_blank"}

```{r}
#| echo: false

# PB neurons celltype80 and 54
annot <- ("celltype54")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "ventral")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype54.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype54.html){target="_blank"}

### Chemoreceptors

```{r}
#| echo: false

# PU neurons celltype13-15
annot <- ("celltype13")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype13.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype13.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype13-15
annot <- ("celltype14")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype14.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype14.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype13-15
annot <- ("celltype15")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype15.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype15.html){target="_blank"}

### Other sensory cells (unclear modality)

```{r}
#| echo: false
#SNgolden
annot <- ("celltype16")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype16.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype16.html){target="_blank"}

```{r}
#| echo: false
#SNhorn
annot <- ("celltype17")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype17.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype17.html){target="_blank"}

```{r}
#| echo: false
#SNhook
annot <- ("celltype18")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype18.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype18.html){target="_blank"}

```{r}
#| echo: false
#SNlasso
annot <- ("celltype20")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype20.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype20.html){target="_blank"}

```{r}
#| echo: false
#SNMIP-vc
annot <- ("celltype24")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype24.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype24.html){target="_blank"}

```{r}
#| echo: false
#SNmus
annot <- ("celltype25")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype25.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype25.html){target="_blank"}

```{r}
#| echo: false
#SNantlerPDF
annot <- ("celltype26")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype26.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype26.html){target="_blank"}

```{r}
#| echo: false
#SNPDF-dc
annot <- ("celltype27")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype27.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype27.html){target="_blank"}

```{r}
#| echo: false
#	SN0DLSO1.2-4
annot <- ("celltype28")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype28.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype28.html){target="_blank"}

```{r}
#| echo: false
#	SN0DLSO1.2-3
annot <- ("celltype29")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype29.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype29.html){target="_blank"}

```{r}
#| echo: false
#	SN0DLSO1.2-1
annot <- ("celltype30")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype30.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype30.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO2
annot <- ("celltype31")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype31.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype31.html){target="_blank"}

```{r}
#| echo: false
#	SNMB2
annot <- ("celltype32")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype32.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype32.html){target="_blank"}

```{r}
#| echo: false
#	SN-ASTC
annot <- ("celltype39")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype39.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype39.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO3-PDF
annot <- ("celltype40")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype40.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype40.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO3
annot <- ("celltype41")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype41.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype41.html){target="_blank"}

```{r}
#| echo: false
#	SN-IRP2-burs
annot <- ("celltype43")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype43.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype43.html){target="_blank"}

```{r}
#| echo: false
#	SN-IRP2-FMRF
annot <- ("celltype44")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype44.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype44.html){target="_blank"}

```{r}
#| echo: false
#	SN-MIP4
annot <- ("celltype45")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype45.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype45.html){target="_blank"}

```{r}
#| echo: false
#	SN-MIP1
annot <- ("celltype46")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype46.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype46.html){target="_blank"}

```{r}
#| echo: false
#	SN-WLD
annot <- ("celltype48")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype48.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype48.html){target="_blank"}

```{r}
#| echo: false
#	SN-DSO
annot <- ("celltype49")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype49.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype49.html){target="_blank"}

```{r}
#| echo: false
#	SNbicil
annot <- ("celltype50")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
rgl <- plot_cell_rgl(annot, "Sensory neuron", "anterior")
widgets <- rglwidget(scene3d(rgl))
# make snapshot
path1 <- paste("_site/snapshots/", annot, ".png", sep = "")
rgl.snapshot(path1)
close3d()

path2 <- paste("_site/celltype_RGLs/", annot, ".html", sep = "")
htmlwidgets::saveWidget(widgets, path2)

```

![](snapshots/celltype50.png){height=200}

[3D visualisation](_site/celltype_RGLs/celltype50.html){target="_blank"}