### 2024-10-21 supplements/celltype_compendium_website/index.qmd
cell type compendium: clean up

---
title: "*Platynereis dumerilii* three-day-old connectome --- cell-type compendium"
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
---

```{r}
#| echo: false
#| include: FALSE
library(dplyr)
library(rio)
library(tinytable)
library(knitr)
library(gt)
library(catmaid)
library(rgl)
library(htmltools)

# catmaid connection
#conn_http1 <- catmaid_login(
#  server="https://catmaid.jekelylab.ex.ac.uk/", 
#  authname="AnonymousUser",
#  config=httr::config(ssl_verifypeer=0, http_version=1)
#)

# catmaid connection, needs username, password AND token - weird!
{
  # can run this separate file using source function
  conn <- source("~/R/conn.R")
  #for larger calls we need to use http/1, see https://www.gitmemory.com/issue/natverse/rcatmaid/158/641537466
  #for this we configure to http/1.1
#  conn_http1 = catmaid_login(conn=conn, config=httr::config(ssl_verifypeer=0, http_version=1))
}
```

# Neuronal cell types

## Sensory neurons

```{r}
#| echo: false

celltypes <- rio::import("../Supplementary_Table1.csv")

#functions for table and neuron plot
plot_cell_name <- function(annotation, type_of_cell){
  name <- celltypes |>
  filter(`Sensory/inter/motor neuron` == type_of_cell) |> 
  filter(`CATMAID annotation` == annotation) |> 
  pull(`CATMAID annotation`, `name of cell type`)

kable(name, col.names = c("name of cell type", "CATMAID annotation"))
}

plot_cell_chars <- function(annotation, type_of_cell){
  celltypes |>
  filter(`Sensory/inter/motor neuron` == type_of_cell) |>
  filter(`CATMAID annotation` == annotation) |>   
  select(`number of cells`, `soma position`, `region`, `transmitter phenotype`, `postsynaptic partners (<10 synapses)`, `presynaptic partners (<10 synapses)`, `reference`) |>
  kable()
}

```

### Photoreceptors

#### Rhabdomeric photoreceptors

```{r}
#| echo: false
# PRC
annot <- "celltype1"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```
![](snapshots/celltype1.png){height=200}

[3D visualisation](celltype_RGLs/celltype1.html){target="_blank"}

```{r}
#| echo: false

# eyespot-PRCR3
annot <- "celltype33"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype33.png){height=200}

[3D visualisation](celltype_RGLs/celltype33.html){target="_blank"}

```{r}
#| echo: false

# eyespot-PRCR1
annot <- "celltype34"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype34.png){height=200}

[3D visualisation](celltype_RGLs/celltype34.html){target="_blank"}

#### Ciliary photoreceptors {#sec-cPRC}

```{r}
#| echo: false

# cPRC
annot <- "celltype5"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype5.png){height=200}

[3D visualisation](celltype_RGLs/celltype5.html){target="_blank"}

### Mechanoreceptors

#### CR (collar receptor) hydrodynamic vibration-receptor neurons

```{r}
#| echo: false

# hCR
annot <- "celltype100"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype100.png){height=200}

[3D visualisation](celltype_RGLs/celltype100.html){target="_blank"}

```{r}
#| echo: false

# ventralpygCR
annot <- "celltype101"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype101.png){height=200}

[3D visualisation](celltype_RGLs/celltype101.html){target="_blank"}

```{r}
#| echo: false

# doCRunp
annot <- "celltype102"
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype102.png){height=200}

[3D visualisation](celltype_RGLs/celltype102.html){target="_blank"}

#### PU (penetrating uniciliated) mechanosensory neurons

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype88")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype88.png){height=200}

[3D visualisation](celltype_RGLs/celltype88.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype89")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype89.png){height=200}

[3D visualisation](celltype_RGLs/celltype89.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
#VentraltrunkPUunp
annot <- ("celltype90")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype90.png){height=200}

[3D visualisation](celltype_RGLs/celltype90.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype91")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype91.png){height=200}

[3D visualisation](celltype_RGLs/celltype91.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
#hPUc1
annot <- ("celltype92")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype92.png){height=200}

[3D visualisation](celltype_RGLs/celltype92.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype93")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype93.png){height=200}

[3D visualisation](celltype_RGLs/celltype93.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype94")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype94.png){height=200}

[3D visualisation](celltype_RGLs/celltype94.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype95")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype95.png){height=200}

[3D visualisation](celltype_RGLs/celltype95.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype96")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype96.png){height=200}

[3D visualisation](celltype_RGLs/celltype96.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype97")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype97.png){height=200}

[3D visualisation](celltype_RGLs/celltype97.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype88-98
annot <- ("celltype98")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype98.png){height=200}

[3D visualisation](celltype_RGLs/celltype98.html){target="_blank"}

#### PB (penetrating biciliated) mechanosensory neurons  {#sec-PB}

```{r}
#| echo: false

# PB neurons celltype80 and 54
annot <- ("celltype80")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype80.png){height=200}

[3D visualisation](celltype_RGLs/celltype80.html){target="_blank"}

```{r}
#| echo: false

# PB neurons celltype80 and 54
#pygPBunp
annot <- ("celltype54")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype54.png){height=200}

[3D visualisation](celltype_RGLs/celltype54.html){target="_blank"}

#### Flow-sensory MS neurons

```{r}
#| echo: false

# MS neurons celltype35-38
annot <- ("celltype35")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype35.png){height=200}

[3D visualisation](celltype_RGLs/celltype35.html){target="_blank"}

```{r}
#| echo: false

# MS neurons celltype35-38
annot <- ("celltype36")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype36.png){height=200}

[3D visualisation](celltype_RGLs/celltype36.html){target="_blank"}

```{r}
#| echo: false

# MS neurons celltype35-38
annot <- ("celltype37")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype37.png){height=200}

[3D visualisation](celltype_RGLs/celltype37.html){target="_blank"}

```{r}
#| echo: false

# MS neurons celltype35-38
annot <- ("celltype38")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype38.png){height=200}

[3D visualisation](celltype_RGLs/celltype38.html){target="_blank"}

#### Chaetal mechanosensory neurons

```{r}
#| echo: false

# chaeMech
annot <- ("celltype71")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype71.png){height=200}

[3D visualisation](celltype_RGLs/celltype71.html){target="_blank"}

#### Other putative mechanosensory neurons in the mechanosensory girdle

```{r}
#| echo: false
#SNantlerPDF
annot <- ("celltype26")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype26.png){height=200}

[3D visualisation](celltype_RGLs/celltype26.html){target="_blank"}

```{r}
#| echo: false
#SNblunt
annot <- ("celltype148")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype148.png){height=200}

[3D visualisation](celltype_RGLs/celltype148.html){target="_blank"}

```{r}
#| echo: false
#SNbronto
annot <- ("celltype168")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype168.png){height=200}

[3D visualisation](celltype_RGLs/celltype168.html){target="_blank"}

```{r}
#| echo: false
#SNFVa
annot <- ("celltype143")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype143.png){height=200}

[3D visualisation](celltype_RGLs/celltype143.html){target="_blank"}

```{r}
#| echo: false
#SNPDF-pyg
annot <- ("celltype115")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype115.png){height=200}

[3D visualisation](celltype_RGLs/celltype115.html){target="_blank"}

```{r}
#| echo: false
#SNpygM
annot <- ("celltype170")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype170.png){height=200}

[3D visualisation](celltype_RGLs/celltype170.html){target="_blank"}

### Nuchal organ chemoreceptors

```{r}
#| echo: false

# PU neurons celltype13-15
annot <- ("celltype13")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype13.png){height=200}

[3D visualisation](celltype_RGLs/celltype13.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype13-15
annot <- ("celltype14")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype14.png){height=200}

[3D visualisation](celltype_RGLs/celltype14.html){target="_blank"}

```{r}
#| echo: false

# PU neurons celltype13-15
annot <- ("celltype15")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype15.png){height=200}

[3D visualisation](celltype_RGLs/celltype15.html){target="_blank"}

### Mushroom body sensory cells

```{r}
#| echo: false
#SNtrpa
annot <- ("celltype190")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype190.png){height=200}

[3D visualisation](celltype_RGLs/celltype190.html){target="_blank"}

```{r}
#| echo: false
#SNtorii
annot <- ("celltype187")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype187.png){height=200}

[3D visualisation](celltype_RGLs/celltype187.html){target="_blank"}

```{r}
#| echo: false
#SNmus
annot <- ("celltype25")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype25.png){height=200}

[3D visualisation](celltype_RGLs/celltype25.html){target="_blank"}

```{r}
#| echo: false
#	SNMB2
annot <- ("celltype32")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype32.png){height=200}

[3D visualisation](celltype_RGLs/celltype32.html){target="_blank"}

```{r}
#| echo: false
#SNlasso
annot <- ("celltype20")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype20.png){height=200}

[3D visualisation](celltype_RGLs/celltype20.html){target="_blank"}

```{r}
#| echo: false
#SNhorn
annot <- ("celltype17")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype17.png){height=200}

[3D visualisation](celltype_RGLs/celltype17.html){target="_blank"}

```{r}
#| echo: false
#SNhook
annot <- ("celltype18")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype18.png){height=200}

[3D visualisation](celltype_RGLs/celltype18.html){target="_blank"}

```{r}
#| echo: false
#SNgolden
annot <- ("celltype16")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype16.png){height=200}

[3D visualisation](celltype_RGLs/celltype16.html){target="_blank"}

```{r}
#| echo: false
#SN-NS27
annot <- ("celltype137")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype137.png){height=200}

[3D visualisation](celltype_RGLs/celltype137.html){target="_blank"}

```{r}
#| echo: false
#SN-NS19
annot <- ("celltype133")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype133.png){height=200}

[3D visualisation](celltype_RGLs/celltype133.html){target="_blank"}

```{r}
#| echo: false
#SN-NS18
annot <- ("celltype138")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype138.png){height=200}

[3D visualisation](celltype_RGLs/celltype138.html){target="_blank"}

```{r}
#| echo: false
#SN-NS17
annot <- ("celltype131")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype131.png){height=200}

[3D visualisation](celltype_RGLs/celltype131.html){target="_blank"}

```{r}
#| echo: false
#SN-NS1
annot <- ("celltype110")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype110.png){height=200}

[3D visualisation](celltype_RGLs/celltype110.html){target="_blank"}

### Apical organ sensory cells

```{r}
#| echo: false
#	SNadNS22
annot <- ("celltype141")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype141.png){height=200}

[3D visualisation](celltype_RGLs/celltype141.html){target="_blank"}

```{r}
#| echo: false
#	SN47Ach
annot <- ("celltype52")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype52.png){height=200}

[3D visualisation](celltype_RGLs/celltype52.html){target="_blank"}

```{r}
#| echo: false
#	SN-YF5cil
annot <- ("celltype132")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype132.png){height=200}

[3D visualisation](celltype_RGLs/celltype132.html){target="_blank"}

```{r}
#| echo: false
#	SN47Ach
annot <- ("celltype52")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype52.png){height=200}

[3D visualisation](celltype_RGLs/celltype52.html){target="_blank"}

```{r}
#| echo: false
#	SN-WLD
annot <- ("celltype48")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype48.png){height=200}

[3D visualisation](celltype_RGLs/celltype48.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS6
annot <- ("celltype135")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype135.png){height=200}

[3D visualisation](celltype_RGLs/celltype135.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS5
annot <- ("celltype130")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype130.png){height=200}

[3D visualisation](celltype_RGLs/celltype130.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS4
annot <- ("celltype112")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype112.png){height=200}

[3D visualisation](celltype_RGLs/celltype112.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS3
annot <- ("celltype111")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype111.png){height=200}

[3D visualisation](celltype_RGLs/celltype111.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS29
annot <- ("celltype136")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype136.png){height=200}

[3D visualisation](celltype_RGLs/celltype136.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS22
annot <- ("celltype129")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype129.png){height=200}

[3D visualisation](celltype_RGLs/celltype129.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS20
annot <- ("celltype134")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype134.png){height=200}

[3D visualisation](celltype_RGLs/celltype134.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS16
annot <- ("celltype139")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype139.png){height=200}

[3D visualisation](celltype_RGLs/celltype139.html){target="_blank"}

```{r}
#| echo: false
#	SN-NS15
annot <- ("celltype113")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype113.png){height=200}

[3D visualisation](celltype_RGLs/celltype113.html){target="_blank"}

```{r}
#| echo: false
#	SN-MIP4
annot <- ("celltype45")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype45.png){height=200}

[3D visualisation](celltype_RGLs/celltype45.html){target="_blank"}

```{r}
#| echo: false
#	SN-MIP1
annot <- ("celltype46")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype46.png){height=200}

[3D visualisation](celltype_RGLs/celltype46.html){target="_blank"}

```{r}
#| echo: false
#	SN-IRP2-FMRF
annot <- ("celltype44")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype44.png){height=200}

[3D visualisation](celltype_RGLs/celltype44.html){target="_blank"}

```{r}
#| echo: false
#	SN-IRP2-burs
annot <- ("celltype43")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype43.png){height=200}

[3D visualisation](celltype_RGLs/celltype43.html){target="_blank"}

```{r}
#| echo: false
#	SN-ASTC
annot <- ("celltype39")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype39.png){height=200}

[3D visualisation](celltype_RGLs/celltype39.html){target="_blank"}

See also @sec-cPRC for the apical organ ciliary photoreceptors.

### Dorsolateral sensory organ sensory cells

```{r}
#| echo: false
#	SN0DLSO1.2-4
annot <- ("celltype28")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype28.png){height=200}

[3D visualisation](celltype_RGLs/celltype28.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO3-PDF
annot <- ("celltype40")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype40.png){height=200}

[3D visualisation](celltype_RGLs/celltype40.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO3
annot <- ("celltype41")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype41.png){height=200}

[3D visualisation](celltype_RGLs/celltype41.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO2
annot <- ("celltype31")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype31.png){height=200}

[3D visualisation](celltype_RGLs/celltype31.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO1.3
annot <- ("celltype163")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype163.png){height=200}

[3D visualisation](celltype_RGLs/celltype163.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO1.2-3
annot <- ("celltype29")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype29.png){height=200}

[3D visualisation](celltype_RGLs/celltype29.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO1.2-1
annot <- ("celltype30")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype30.png){height=200}

[3D visualisation](celltype_RGLs/celltype30.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO1.1NP
annot <- ("celltype172")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype172.png){height=200}

[3D visualisation](celltype_RGLs/celltype172.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO1.1
annot <- ("celltype173")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype173.png){height=200}

[3D visualisation](celltype_RGLs/celltype173.html){target="_blank"}

```{r}
#| echo: false
#	SN-DLSO1.0
annot <- ("celltype103")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype103.png){height=200}

[3D visualisation](celltype_RGLs/celltype103.html){target="_blank"}

### Other sensory cells (unclear modality)

```{r}
#| echo: false
#SNMIP-vc
annot <- ("celltype24")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype24.png){height=200}

[3D visualisation](celltype_RGLs/celltype24.html){target="_blank"}

```{r}
#| echo: false
#SNPDF-dc
annot <- ("celltype27")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype27.png){height=200}

[3D visualisation](celltype_RGLs/celltype27.html){target="_blank"}

```{r}
#| echo: false
#	SN-DSO
annot <- ("celltype49")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype49.png){height=200}

[3D visualisation](celltype_RGLs/celltype49.html){target="_blank"}

```{r}
#| echo: false
#	SNbicil
annot <- ("celltype50")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype50.png){height=200}

[3D visualisation](celltype_RGLs/celltype50.html){target="_blank"}

```{r}
#| echo: false
#SNasym
annot <- ("celltype51")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype51.png){height=200}

[3D visualisation](celltype_RGLs/celltype51.html){target="_blank"}

```{r}
#| echo: false
#SNaanc
annot <- ("celltype114")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype114.png){height=200}

[3D visualisation](celltype_RGLs/celltype114.html){target="_blank"}

```{r}
#| echo: false
#SNstiff
annot <- ("celltype167")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")

```

![](snapshots/celltype167.png){height=200}

[3D visualisation](celltype_RGLs/celltype167.html){target="_blank"}

## Interneurons

### Visual system interneurons

```{r}
#| echo: false
#IN1
annot <- ("celltype2")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype2.png){height=200}
[3D visualisation](celltype_RGLs/celltype2.html){target="_blank"}

```{r}
#| echo: false
#INton
annot <- ("celltype3")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype3.png){height=200}
[3D visualisation](celltype_RGLs/celltype3.html){target="_blank"}

```{r}
#| echo: false
#INR1
annot <- ("celltype171")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype171.png){height=200}
[3D visualisation](celltype_RGLs/celltype171.html){target="_blank"}

### Ciliomotor interneurons

```{r}
#| echo: false
#cMNATO
annot <- ("celltype11")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype11.png){height=200}

[3D visualisation](celltype_RGLs/celltype11.html){target="_blank"}

```{r}
#| echo: false
#cMNdc
annot <- ("celltype12")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype12.png){height=200}

[3D visualisation](celltype_RGLs/celltype12.html){target="_blank"}

```{r}
#| echo: false
#cMNPDF-vcl1
annot <- ("celltype10")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype10.png){height=200}

[3D visualisation](celltype_RGLs/celltype10.html){target="_blank"}

### Apical organ interneurons

```{r}
#| echo: false
#INNOS
annot <- ("celltype7")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype7.png){height=200}
[3D visualisation](celltype_RGLs/celltype7.html){target="_blank"}

```{r}
#| echo: false
#INRGWa
annot <- ("celltype6")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype6.png){height=200}
[3D visualisation](celltype_RGLs/celltype6.html){target="_blank"}

### Central brain interneurons

```{r}
#| echo: false
#INexsn
annot <- ("celltype188")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype188.png){height=200}
[3D visualisation](celltype_RGLs/celltype188.html){target="_blank"}

```{r}
#| echo: false
#INface
annot <- ("celltype193")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype193.png){height=200}
[3D visualisation](celltype_RGLs/celltype193.html){target="_blank"}

```{r}
#| echo: false
#INmidL
annot <- ("celltype118")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype118.png){height=200}
[3D visualisation](celltype_RGLs/celltype118.html){target="_blank"}

```{r}
#| echo: false
#INmus
annot <- ("celltype174")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype174.png){height=200}
[3D visualisation](celltype_RGLs/celltype174.html){target="_blank"}

```{r}
#| echo: false
#INpara
annot <- ("celltype196")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype196.png){height=200}
[3D visualisation](celltype_RGLs/celltype196.html){target="_blank"}

```{r}
#| echo: false
#INsn
annot <- ("celltype57")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype57.png){height=200}
[3D visualisation](celltype_RGLs/celltype57.html){target="_blank"}

```{r}
#| echo: false
#INsqNSasym
annot <- ("celltype179")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype179.png){height=200}
[3D visualisation](celltype_RGLs/celltype179.html){target="_blank"}

### Central brain decussating interneurons

```{r}
#| echo: false
#INarc1
annot <- ("celltype55")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype55.png){height=200}
[3D visualisation](celltype_RGLs/celltype55.html){target="_blank"}

```{r}
#| echo: false
#INarc2
annot <- ("celltype56")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype56.png){height=200}
[3D visualisation](celltype_RGLs/celltype56.html){target="_blank"}

```{r}
#| echo: false
#INcross
annot <- ("celltype123")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype123.png){height=200}
[3D visualisation](celltype_RGLs/celltype123.html){target="_blank"}

```{r}
#| echo: false
#INdecussfoot
annot <- ("celltype152")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype152.png){height=200}
[3D visualisation](celltype_RGLs/celltype152.html){target="_blank"}

```{r}
#| echo: false
#INdecusshook
annot <- ("celltype153")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype153.png){height=200}
[3D visualisation](celltype_RGLs/celltype153.html){target="_blank"}

```{r}
#| echo: false
#INdecussPre
annot <- ("celltype22")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype22.png){height=200}
[3D visualisation](celltype_RGLs/celltype22.html){target="_blank"}

```{r}
#| echo: false
#INpro
annot <- ("celltype124")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype124.png){height=200}
[3D visualisation](celltype_RGLs/celltype124.html){target="_blank"}

```{r}
#| echo: false
#INproT2
annot <- ("celltype125")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype125.png){height=200}
[3D visualisation](celltype_RGLs/celltype125.html){target="_blank"}

```{r}
#| echo: false
#INW
annot <- ("celltype117")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype117.png){height=200}
[3D visualisation](celltype_RGLs/celltype117.html){target="_blank"}

### Central brain descending interneurons

```{r}
#| echo: false
#INbiax
annot <- ("celltype76")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype76.png){height=200}
[3D visualisation](celltype_RGLs/celltype76.html){target="_blank"}

```{r}
#| echo: false
#INdescLuqinPDF
annot <- ("celltype175")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype175.png){height=200}
[3D visualisation](celltype_RGLs/celltype175.html){target="_blank"}

```{r}
#| echo: false
#INUturn
annot <- ("celltype144")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype144.png){height=200}
[3D visualisation](celltype_RGLs/celltype144.html){target="_blank"}

### Other brain decussating interneurons

```{r}
#| echo: false
#INdecussM
annot <- ("celltype199")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype199.png){height=200}
[3D visualisation](celltype_RGLs/celltype199.html){target="_blank"}

### Mushroom body intrinsic interneurons

```{r}
#| echo: false
#INMBtype6
annot <- ("celltype120")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype120.png){height=200}
[3D visualisation](celltype_RGLs/celltype120.html){target="_blank"}

```{r}
#| echo: false
#INMBtype5
annot <- ("celltype121")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype121.png){height=200}
[3D visualisation](celltype_RGLs/celltype121.html){target="_blank"}

```{r}
#| echo: false
#INMBtype7
annot <- ("celltype122")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype122.png){height=200}
[3D visualisation](celltype_RGLs/celltype122.html){target="_blank"}

```{r}
#| echo: false
#INMBtype9
annot <- ("celltype197")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype197.png){height=200}
[3D visualisation](celltype_RGLs/celltype197.html){target="_blank"}

```{r}
#| echo: false
#INMBtype2
annot <- ("celltype198")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype198.png){height=200}
[3D visualisation](celltype_RGLs/celltype198.html){target="_blank"}

### Mushroom body projection neurons

```{r}
#| echo: false
#MBmouth
annot <- ("celltype189")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype189.png){height=200}
[3D visualisation](celltype_RGLs/celltype189.html){target="_blank"}

```{r}
#| echo: false
#INMBPDF
annot <- ("celltype184")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype184.png){height=200}
[3D visualisation](celltype_RGLs/celltype184.html){target="_blank"}

```{r}
#| echo: false
#INbigloop
annot <- ("celltype140")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype140.png){height=200}
[3D visualisation](celltype_RGLs/celltype140.html){target="_blank"}

```{r}
#| echo: false
#INhorn
annot <- ("celltype183")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype183.png){height=200}
[3D visualisation](celltype_RGLs/celltype183.html){target="_blank"}

```{r}
#| echo: false
#INMBdescFMRF
annot <- ("celltype185")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype185.png){height=200}
[3D visualisation](celltype_RGLs/celltype185.html){target="_blank"}

```{r}
#| echo: false
#INtorii
annot <- ("celltype191")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype191.png){height=200}
[3D visualisation](celltype_RGLs/celltype191.html){target="_blank"}

```{r}
#| echo: false
#INUturnMB
annot <- ("celltype192")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype192.png){height=200}
[3D visualisation](celltype_RGLs/celltype192.html){target="_blank"}

```{r}
#| echo: false
#INMBdesc3
annot <- ("celltype194")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype194.png){height=200}
[3D visualisation](celltype_RGLs/celltype194.html){target="_blank"}

```{r}
#| echo: false
#INMBdesc2
annot <- ("celltype195")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype195.png){height=200}
[3D visualisation](celltype_RGLs/celltype195.html){target="_blank"}

```{r}
#| echo: false
#INrope
annot <- ("celltype58")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype58.png){height=200}
[3D visualisation](celltype_RGLs/celltype58.html){target="_blank"}

### Head dorsolateral interneuron

```{r}
#| echo: false
#INpreSer
annot <- ("celltype4")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype4.png){height=200}
[3D visualisation](celltype_RGLs/celltype4.html){target="_blank"}

```{r}
#| echo: false
#INlasso
annot <- ("celltype21")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype21.png){height=200}
[3D visualisation](celltype_RGLs/celltype21.html){target="_blank"}

```{r}
#| echo: false
#INpreMN
annot <- ("celltype23")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype23.png){height=200}
[3D visualisation](celltype_RGLs/celltype23.html){target="_blank"}

```{r}
#| echo: false
#INdc
annot <- ("celltype104")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype104.png){height=200}
[3D visualisation](celltype_RGLs/celltype104.html){target="_blank"}

```{r}
#| echo: false
#INfoot
annot <- ("celltype116")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype116.png){height=200}
[3D visualisation](celltype_RGLs/celltype116.html){target="_blank"}

```{r}
#| echo: false
#INhook
annot <- ("celltype119")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype119.png){height=200}
[3D visualisation](celltype_RGLs/celltype119.html){target="_blank"}

```{r}
#| echo: false
#INpear
annot <- ("celltype126")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype126.png){height=200}
[3D visualisation](celltype_RGLs/celltype126.html){target="_blank"}

```{r}
#| echo: false
#INZ
annot <- ("celltype127")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype127.png){height=200}
[3D visualisation](celltype_RGLs/celltype127.html){target="_blank"}

```{r}
#| echo: false
#	INlasso-postSN
annot <- ("celltype160")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype160.png){height=200}
[3D visualisation](celltype_RGLs/celltype160.html){target="_blank"}

```{r}
#| echo: false
#INDLSO
annot <- ("celltype186")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype186.png){height=200}
[3D visualisation](celltype_RGLs/celltype186.html){target="_blank"}

```{r}
#| echo: false
#INmantis
annot <- ("celltype201")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype201.png){height=200}
[3D visualisation](celltype_RGLs/celltype201.html){target="_blank"}

```{r}
#| echo: false
#MC2biax-1
annot <- ("celltype42")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype42.png){height=200}
[3D visualisation](celltype_RGLs/celltype42.html){target="_blank"}

```{r}
#| echo: false
#INCrossbow
annot <- ("celltype53")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype53.png){height=200}
[3D visualisation](celltype_RGLs/celltype53.html){target="_blank"}

### Head, mechanosensory girdle interneurons

```{r}
#| echo: false
#INsplitPUh
annot <- ("celltype83")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype83.png){height=200}
[3D visualisation](celltype_RGLs/celltype83.html){target="_blank"}

```{r}
#| echo: false
#INsplitSpin
annot <- ("celltype202")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype202.png){height=200}
[3D visualisation](celltype_RGLs/celltype202.html){target="_blank"}

```{r}
#| echo: false
#INMC3
annot <- ("celltype75")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype75.png){height=200}
[3D visualisation](celltype_RGLs/celltype75.html){target="_blank"}

### Ventral nerve cord, mechanosensory girdle interneurons

```{r}
#| echo: false
#INCM
annot <- ("celltype60")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype60.png){height=200}
[3D visualisation](celltype_RGLs/celltype60.html){target="_blank"}

```{r}
#| echo: false
#INsplitCR
annot <- ("celltype73")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype73.png){height=200}
[3D visualisation](celltype_RGLs/celltype73.html){target="_blank"}

```{r}
#| echo: false
#INsplitCRATO
annot <- ("celltype74")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype74.png){height=200}
[3D visualisation](celltype_RGLs/celltype74.html){target="_blank"}

```{r}
#| echo: false
#INsplitPB-RF/Ya
annot <- ("celltype77")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype77.png){height=200}
[3D visualisation](celltype_RGLs/celltype77.html){target="_blank"}

```{r}
#| echo: false
#INsplitPBant
annot <- ("celltype78")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype78.png){height=200}
[3D visualisation](celltype_RGLs/celltype78.html){target="_blank"}

```{r}
#| echo: false
#INsplitPB
annot <- ("celltype79")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype79.png){height=200}
[3D visualisation](celltype_RGLs/celltype79.html){target="_blank"}

```{r}
#| echo: false
#INsplitBronto
annot <- ("celltype149")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype149.png){height=200}
[3D visualisation](celltype_RGLs/celltype149.html){target="_blank"}

```{r}
#| echo: false
#INsplitVent
annot <- ("celltype157")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype157.png){height=200}
[3D visualisation](celltype_RGLs/celltype157.html){target="_blank"}

### Pygidial interneurons

```{r}
#| echo: false
#INasc-pyg
annot <- ("celltype147")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype147.png){height=200}
[3D visualisation](celltype_RGLs/celltype147.html){target="_blank"}

```{r}
#| echo: false
#INFVa-pyg 
annot <- ("celltype72")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype72.png){height=200}
[3D visualisation](celltype_RGLs/celltype72.html){target="_blank"}

```{r}
#| echo: false
#INATOpyg
annot <- ("celltype145")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype145.png){height=200}
[3D visualisation](celltype_RGLs/celltype145.html){target="_blank"}

### Peripheral interneurons

```{r}
#| echo: false
#INbackcross
annot <- ("celltype105")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype105.png){height=200}
[3D visualisation](celltype_RGLs/celltype105.html){target="_blank"}

### Other trunk interneurons

```{r}
#| echo: false
#INchaeMech
annot <- ("celltype70")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype70.png){height=200}
[3D visualisation](celltype_RGLs/celltype70.html){target="_blank"}

```{r}
#| echo: false
#INbiaxLeg
annot <- ("celltype108")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype108.png){height=200}
[3D visualisation](celltype_RGLs/celltype108.html){target="_blank"}

```{r}
#| echo: false
#INbiaxH
annot <- ("celltype109")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype109.png){height=200}
[3D visualisation](celltype_RGLs/celltype109.html){target="_blank"}

```{r}
#| echo: false
#INcomm-Upstairs
annot <- ("celltype146")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype146.png){height=200}
[3D visualisation](celltype_RGLs/celltype146.html){target="_blank"}

```{r}
#| echo: false
#INleucoPU
annot <- ("celltype150")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype150.png){height=200}
[3D visualisation](celltype_RGLs/celltype150.html){target="_blank"}

```{r}
#| echo: false
#INpreLadder
annot <- ("celltype154")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype154.png){height=200}
[3D visualisation](celltype_RGLs/celltype154.html){target="_blank"}

```{r}
#| echo: false
#INcomm-Upcross
annot <- ("celltype155")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype155.png){height=200}
[3D visualisation](celltype_RGLs/celltype155.html){target="_blank"}

```{r}
#| echo: false
#INcomm-DownL
annot <- ("celltype158")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype158.png){height=200}
[3D visualisation](celltype_RGLs/celltype158.html){target="_blank"}

```{r}
#| echo: false
#INcommascFV
annot <- ("celltype169")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype169.png){height=200}
[3D visualisation](celltype_RGLs/celltype169.html){target="_blank"}

```{r}
#| echo: false
#INcosplit
annot <- ("celltype176")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype176.png){height=200}
[3D visualisation](celltype_RGLs/celltype176.html){target="_blank"}

```{r}
#| echo: false
#INcommdescFVa
annot <- ("celltype177")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype177.png){height=200}
[3D visualisation](celltype_RGLs/celltype177.html){target="_blank"}

```{r}
#| echo: false
#INbiaxHmid
annot <- ("celltype180")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype180.png){height=200}
[3D visualisation](celltype_RGLs/celltype180.html){target="_blank"}

```{r}
#| echo: false
#INiaStiff
annot <- ("celltype200")
plot_cell_name(annot, "Interneuron")
plot_cell_chars(annot, "Interneuron")

```

![](snapshots/celltype200.png){height=200}
[3D visualisation](celltype_RGLs/celltype200.html){target="_blank"}

## Motor neurons

### Ciliomotor neurons

```{r}
#| echo: false
#MC
annot <- ("celltype9")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype9.png){height=200}
[3D visualisation](celltype_RGLs/celltype9.html){target="_blank"}

```{r}
#| echo: false
#MNakro
annot <- ("celltype164")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype164.png){height=200}
[3D visualisation](celltype_RGLs/celltype164.html){target="_blank"}

```{r}
#| echo: false
#	Ser-h1
annot <- ("celltype8")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype8.png){height=200}
[3D visualisation](celltype_RGLs/celltype8.html){target="_blank"}

```{r}
#| echo: false
#	MNant
annot <- ("celltype19")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype19.png){height=200}
[3D visualisation](celltype_RGLs/celltype19.html){target="_blank"}

```{r}
#| echo: false
#	Ser-tr1
annot <- ("celltype142")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype142.png){height=200}
[3D visualisation](celltype_RGLs/celltype142.html){target="_blank"}

```{r}
#| echo: false
#Loop
annot <- ("celltype59")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype59.png){height=200}
[3D visualisation](celltype_RGLs/celltype59.html){target="_blank"}

```{r}
#| echo: false
#	MN3
annot <- ("celltype84")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype84.png){height=200}
[3D visualisation](celltype_RGLs/celltype84.html){target="_blank"}

### Mixed ciliomotor-musclemotor neurons

```{r}
#| echo: false
#	MN1
annot <- ("celltype85")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype85.png){height=200}
[3D visualisation](celltype_RGLs/celltype85.html){target="_blank"}

```{r}
#| echo: false
#	MN2
annot <- ("celltype86")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype86.png){height=200}
[3D visualisation](celltype_RGLs/celltype86.html){target="_blank"}

### Other head motoneurons

```{r}
#| echo: false
#	MNheadV
annot <- ("celltype178")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype178.png){height=200}
[3D visualisation](celltype_RGLs/celltype178.html){target="_blank"}

### Glandmotor neurons

```{r}
#| echo: false
#	MNgland-head
annot <- ("celltype166")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype166.png){height=200}
[3D visualisation](celltype_RGLs/celltype166.html){target="_blank"}

```{r}
#| echo: false
#	MNspinning
annot <- ("celltype47")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype47.png){height=200}
[3D visualisation](celltype_RGLs/celltype47.html){target="_blank"}

### Pigment-cell-motor neurons

```{r}
#| echo: false
#	cioMNcover
annot <- ("celltype82")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype82.png){height=200}
[3D visualisation](celltype_RGLs/celltype82.html){target="_blank"}

```{r}
#| echo: false
#	MC3cover
annot <- ("celltype87")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype87.png){height=200}
[3D visualisation](celltype_RGLs/celltype87.html){target="_blank"}

```{r}
#| echo: false
#	pygPBunp
annot <- ("celltype54")
plot_cell_name(annot, "Sensory neuron")
plot_cell_chars(annot, "Sensory neuron")
```

![](snapshots/celltype54.png){height=200}
[3D visualisation](celltype_RGLs/celltype54.html){target="_blank"}

See also @sec-PB for pygPBunp as penetrating biciliated sensory neuron.

### Musclemotor neurons

```{r}
#| echo: false
#	MNspider-ant
annot <- ("celltype61")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype61.png){height=200}
[3D visualisation](celltype_RGLs/celltype61.html){target="_blank"}

```{r}
#| echo: false
#	MNspider-post
annot <- ("celltype62")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype62.png){height=200}
[3D visualisation](celltype_RGLs/celltype62.html){target="_blank"}

```{r}
#| echo: false
#	MNbiramous
annot <- ("celltype63")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype63.png){height=200}
[3D visualisation](celltype_RGLs/celltype63.html){target="_blank"}

```{r}
#| echo: false
#	MNcrab
annot <- ("celltype65")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype65.png){height=200}
[3D visualisation](celltype_RGLs/celltype65.html){target="_blank"}

```{r}
#| echo: false
#	MNhose
annot <- ("celltype66")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype66.png){height=200}
[3D visualisation](celltype_RGLs/celltype66.html){target="_blank"}

```{r}
#| echo: false
#	MNbow
annot <- ("celltype67")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype67.png){height=200}
[3D visualisation](celltype_RGLs/celltype67.html){target="_blank"}

```{r}
#| echo: false
#	MNwave
annot <- ("celltype68")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype68.png){height=200}
[3D visualisation](celltype_RGLs/celltype68.html){target="_blank"}

```{r}
#| echo: false
#	MNacicX
annot <- ("celltype69")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype69.png){height=200}
[3D visualisation](celltype_RGLs/celltype69.html){target="_blank"}

```{r}
#| echo: false
#	MNcommUpL
annot <- ("celltype106")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype106.png){height=200}
[3D visualisation](celltype_RGLs/celltype106.html){target="_blank"}

```{r}
#| echo: false
#	MNob-ipsi
annot <- ("celltype107")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype107.png){height=200}
[3D visualisation](celltype_RGLs/celltype107.html){target="_blank"}

```{r}
#| echo: false
#MNpostv
annot <- ("celltype128")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype128.png){height=200}
[3D visualisation](celltype_RGLs/celltype128.html){target="_blank"}

```{r}
#| echo: false
#MNladder
annot <- ("celltype151")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype151.png){height=200}
[3D visualisation](celltype_RGLs/celltype151.html){target="_blank"}

```{r}
#| echo: false
#MNacic
annot <- ("celltype156")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype156.png){height=200}
[3D visualisation](celltype_RGLs/celltype156.html){target="_blank"}

```{r}
#| echo: false
#MNantacic
annot <- ("celltype161")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype161.png){height=200}
[3D visualisation](celltype_RGLs/celltype161.html){target="_blank"}

```{r}
#| echo: false
#MNpostacic
annot <- ("celltype162")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype162.png){height=200}
[3D visualisation](celltype_RGLs/celltype162.html){target="_blank"}

```{r}
#| echo: false
#MNantelope
annot <- ("celltype181")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype181.png){height=200}
[3D visualisation](celltype_RGLs/celltype181.html){target="_blank"}

```{r}
#| echo: false
#MNsmile
annot <- ("celltype182")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype182.png){height=200}
[3D visualisation](celltype_RGLs/celltype182.html){target="_blank"}

```{r}
#| echo: false
#MNarm
annot <- ("celltype159")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype159.png){height=200}
[3D visualisation](celltype_RGLs/celltype159.html){target="_blank"}

```{r}
#| echo: false
#MNchae
annot <- ("celltype165")
plot_cell_name(annot, "Motoneuron")
plot_cell_chars(annot, "Motoneuron")

```

![](snapshots/celltype165.png){height=200}
[3D visualisation](celltype_RGLs/celltype165.html){target="_blank"}

# Non-neuronal cell types

## Gland cells

```{r}
#| echo: false
#spinGland
annot <- ("celltype_non_neuronal7")
plot_cell_name(annot, "gland cell")
plot_cell_chars(annot, "gland cell")

```

![](snapshots/celltype_non_neuronal7.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal7.html){target="_blank"}

```{r}
#| echo: false
#ciliatedGland
annot <- ("celltype_non_neuronal9")
plot_cell_name(annot, "gland cell")
plot_cell_chars(annot, "gland cell")

```

![](snapshots/celltype_non_neuronal9.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal9.html){target="_blank"}

```{r}
#| echo: false
#MVGland
annot <- ("celltype_non_neuronal17")
plot_cell_name(annot, "gland cell")
plot_cell_chars(annot, "gland cell")

```

![](snapshots/celltype_non_neuronal17.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal17.html){target="_blank"}

```{r}
#| echo: false
#HeadGland
annot <- ("celltype_non_neuronal29")
plot_cell_name(annot, "gland cell")
plot_cell_chars(annot, "gland cell")

```

![](snapshots/celltype_non_neuronal29.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal29.html){target="_blank"}

```{r}
#| echo: false
#InterparaGland
annot <- ("celltype_non_neuronal30")
plot_cell_name(annot, "gland cell")
plot_cell_chars(annot, "gland cell")

```

![](snapshots/celltype_non_neuronal30.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal30.html){target="_blank"}

```{r}
#| echo: false
#spinMicroGland
annot <- ("celltype_non_neuronal31")
plot_cell_name(annot, "gland cell")
plot_cell_chars(annot, "gland cell")

```

![](snapshots/celltype_non_neuronal31.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal31.html){target="_blank"}

## Glia cells

```{r}
#| echo: false
#flat-glia
annot <- ("celltype_non_neuronal15")
plot_cell_name(annot, "glia cell")
plot_cell_chars(annot, "glia cell")

```

![](snapshots/celltype_non_neuronal15.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal15.html){target="_blank"}

```{r}
#| echo: false
#ECmidline
annot <- ("celltype_non_neuronal16")
plot_cell_name(annot, "glia cell")
plot_cell_chars(annot, "glia cell")

```

![](snapshots/celltype_non_neuronal16.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal16.html){target="_blank"}

```{r}
#| echo: false
#Glia-pigmented
annot <- ("celltype_non_neuronal34")
plot_cell_name(annot, "glia cell")
plot_cell_chars(annot, "glia cell")

```

![](snapshots/celltype_non_neuronal34.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal34.html){target="_blank"}

## Multiciliated cells

```{r}
#| echo: false
#nuchalCiliaCilia
annot <- ("celltype_non_neuronal4")
plot_cell_name(annot, "multiciliated cell")
plot_cell_chars(annot, "multiciliated cell")

```

![](snapshots/celltype_non_neuronal4.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal4.html){target="_blank"}

```{r}
#| echo: false
#akrotroch
annot <- ("celltype_non_neuronal1")
plot_cell_name(annot, "multiciliated cell")
plot_cell_chars(annot, "multiciliated cell")

```

![](snapshots/celltype_non_neuronal1.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal1.html){target="_blank"}

```{r}
#| echo: false
#crescentcell
annot <- ("celltype_non_neuronal2")
plot_cell_name(annot, "multiciliated cell")
plot_cell_chars(annot, "multiciliated cell")

```

![](snapshots/celltype_non_neuronal2.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal2.html){target="_blank"}

```{r}
#| echo: false
#prototroch
annot <- ("celltype_non_neuronal3")
plot_cell_name(annot, "multiciliated cell")
plot_cell_chars(annot, "multiciliated cell")

```

![](snapshots/celltype_non_neuronal3.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal3.html){target="_blank"}

```{r}
#| echo: false
#metatroch
annot <- ("celltype_non_neuronal5")
plot_cell_name(annot, "multiciliated cell")
plot_cell_chars(annot, "multiciliated cell")

```

![](snapshots/celltype_non_neuronal5.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal5.html){target="_blank"}

```{r}
#| echo: false
#paratroch
annot <- ("celltype_non_neuronal6")
plot_cell_name(annot, "multiciliated cell")
plot_cell_chars(annot, "multiciliated cell")

```

![](snapshots/celltype_non_neuronal6.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal6.html){target="_blank"}

## Pigment cells

```{r}
#| echo: false
#covercell
annot <- ("celltype_non_neuronal8")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal8.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal8.html){target="_blank"}

```{r}
#| echo: false
#eyespot-pigment-cell
annot <- ("celltype_non_neuronal10")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal10.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal10.html){target="_blank"}

```{r}
#| echo: false
#	eye-pigment-cell
annot <- ("celltype_non_neuronal11")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal11.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal11.html){target="_blank"}

```{r}
#| echo: false
#	yolk cover cell 
annot <- ("celltype_non_neuronal14")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal14.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal14.html){target="_blank"}

```{r}
#| echo: false
#	CB-pigment
annot <- ("celltype_non_neuronal32")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal32.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal32.html){target="_blank"}

```{r}
#| echo: false
#	vacuolar-cell-head
annot <- ("celltype_non_neuronal33")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal33.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal33.html){target="_blank"}

```{r}
#| echo: false
#	pygidial-pigment-cell
annot <- ("celltype_non_neuronal35")
plot_cell_name(annot, "pigment cell")
plot_cell_chars(annot, "pigment cell")

```

![](snapshots/celltype_non_neuronal35.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal35.html){target="_blank"}

## Chaetal and acicular cells

```{r}
#| echo: false
#	chaeta
annot <- ("celltype_non_neuronal22")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal22.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal22.html){target="_blank"}

```{r}
#| echo: false
#	acicula
annot <- ("celltype_non_neuronal23")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal23.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal23.html){target="_blank"}

```{r}
#| echo: false
#	acFC-noER
annot <- ("celltype_non_neuronal24")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal24.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal24.html){target="_blank"}

```{r}
#| echo: false
#	chaeFC-hemi
annot <- ("celltype_non_neuronal25")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal25.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal25.html){target="_blank"}

```{r}
#| echo: false
#	chaeFC-ER
annot <- ("celltype_non_neuronal26")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal26.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal26.html){target="_blank"}

```{r}
#| echo: false
#	chaeFC-noER
annot <- ("celltype_non_neuronal27")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal27.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal27.html){target="_blank"}

```{r}
#| echo: false
#	chaeFC-EC
annot <- ("celltype_non_neuronal28")
plot_cell_name(annot, "chaetal complex")
plot_cell_chars(annot, "chaetal complex")

```

![](snapshots/celltype_non_neuronal28.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal28.html){target="_blank"}

## Nephridial cells

```{r}
#| echo: false
#protonephridium
annot <- ("celltype_non_neuronal19")
plot_cell_name(annot, "excretory system")
plot_cell_chars(annot, "excretory system")

```

![](snapshots/celltype_non_neuronal19.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal19.html){target="_blank"}

```{r}
#| echo: false
#nephridium
annot <- ("celltype_non_neuronal20")
plot_cell_name(annot, "excretory system")
plot_cell_chars(annot, "excretory system")

```

![](snapshots/celltype_non_neuronal20.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal20.html){target="_blank"}

```{r}
#| echo: false
#nephridiumTip
annot <- ("celltype_non_neuronal21")
plot_cell_name(annot, "excretory system")
plot_cell_chars(annot, "excretory system")

```

![](snapshots/celltype_non_neuronal21.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal21.html){target="_blank"}

## Muscle cells

```{r}
#| echo: false
annot <- ("celltype_non_neuronal37")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal37.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal37.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal38")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal38.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal38.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal39")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal39.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal39.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal40")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal40.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal40.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal41")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal41.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal41.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal42")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal42.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal42.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal43")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal43.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal43.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal44")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal44.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal44.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal45")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal45.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal45.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal46")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal46.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal46.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal47")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal47.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal47.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal48")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal48.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal48.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal49")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal49.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal49.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal50")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal50.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal50.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal51")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal51.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal51.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal52")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal52.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal52.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal53")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal53.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal53.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal54")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal54.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal54.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal55")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal55.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal55.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal56")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal56.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal56.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal57")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal57.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal57.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal58")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal58.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal58.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal59")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal59.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal59.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal60")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal60.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal60.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal61")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal61.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal61.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal62")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal62.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal62.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal63")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal63.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal63.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal64")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal64.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal64.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal65")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal65.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal65.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal66")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal66.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal66.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal67")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal67.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal67.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal68")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal68.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal68.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal69")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal69.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal69.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal70")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal70.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal70.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal71")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal71.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal71.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal72")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal72.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal72.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal73")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal73.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal73.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal74")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal74.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal74.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal75")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal75.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal75.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal76")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal76.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal76.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal77")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal77.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal77.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal78")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal78.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal78.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal79")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal79.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal79.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal80")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal80.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal80.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal81")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal81.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal81.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal82")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal82.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal82.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal83")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal83.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal83.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal84")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal84.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal84.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal85")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal85.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal85.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal86")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal86.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal86.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal87")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal87.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal87.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal88")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal88.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal88.html){target="_blank"}

```{r}
#| echo: false
annot <- ("celltype_non_neuronal89")
plot_cell_name(annot, "muscle")
plot_cell_chars(annot, "muscle")
```
![](snapshots/celltype_non_neuronal89.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal89.html){target="_blank"}

## Other non-neuronal cell types

```{r}
#| echo: false
#coelothelium
annot <- ("celltype_non_neuronal36")
plot_cell_name(annot, "coelothelium")
plot_cell_chars(annot, "coelothelium")

```

![](snapshots/celltype_non_neuronal36.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal36.html){target="_blank"}

```{r}
#| echo: false
#yolk
annot <- ("celltype_non_neuronal91")
plot_cell_name(annot, "yolk")
plot_cell_chars(annot, "yolk")

```

![](snapshots/celltype_non_neuronal91.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal91.html){target="_blank"}

```{r}
#| echo: false
#meso
annot <- ("celltype_non_neuronal92")
plot_cell_name(annot, "mesoderm")
plot_cell_chars(annot, "mesoderm")

```

![](snapshots/celltype_non_neuronal92.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal92.html){target="_blank"}

```{r}
#| echo: false
#macrophage like
annot <- ("celltype_non_neuronal13")
plot_cell_name(annot, "macrophage-like")
plot_cell_chars(annot, "macrophage-like")

```

![](snapshots/celltype_non_neuronal13.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal13.html){target="_blank"}

```{r}
#| echo: false
#microvillarCell
annot <- ("celltype_non_neuronal18")
plot_cell_name(annot, "microvillar")
plot_cell_chars(annot, "microvillar")

```

![](snapshots/celltype_non_neuronal18.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal18.html){target="_blank"}

```{r}
#| echo: false
#gland-pigment
annot <- ("celltype_non_neuronal12")
plot_cell_name(annot, "gland")
plot_cell_chars(annot, "gland")

```

![](snapshots/celltype_non_neuronal12.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal12.html){target="_blank"}

```{r}
#| echo: false
#EC
annot <- ("celltype_non_neuronal90")
plot_cell_name(annot, "epidermis")
plot_cell_chars(annot, "epidermis")

```

![](snapshots/celltype_non_neuronal90.png){height=200}
[3D visualisation](celltype_RGLs/celltype_non_neuronal90.html){target="_blank"}