### 2025-07-07 supplements/celltype_compendium_website/celltype_compendium/about.qmd
add celltyp ecompendium website and morphological renderings of brain, girdle, MB

---
format:
  html:
    smooth-scroll: true
    theme: [default, custom_index.scss]

---

### Cell type compendium of the three-day-old *Platynereis dumerilii* larva.

<br>

Based on the *Whole-body connectome of a segmented annelid larva*
[connectome paper](https://elifesciences.org/reviewed-preprints/97964v1){.external target="_blank"}.
<br>

#### by the Jékely lab, [Centre for Organismal Studies](https://www.cos.uni-heidelberg.de/en/research-groups/gaspar-jekely){.external target="_blank"}, Heidelberg University