### 2025-07-07 supplements/celltype_compendium_website/index.qmd

---
format:
  html:
    smooth-scroll: true
    theme: [default, custom_index.scss]

---

## Whole-body connectome of a segmented annelid larva

#### Interactive visualisations and resources of the *Platynereis dumerilii* three-day-old larvae whole-body connectome based on the [connectome paper](https://elifesciences.org/reviewed-preprints/97964v1){.external target="_blank"}.

#### by the Jékely lab, [Centre for Organismal Studies](https://www.cos.uni-heidelberg.de/en/research-groups/gaspar-jekely){.external target="_blank"}, Heidelberg University

<br>

::: columns
::: {.column width="31%"}

![](Platynereis_SEM_3d_2.png)
:::
::: {.column width="4%"}
:::
::: {.column width="65%"}

![](network_with_IO_layout.png)
:::
:::

<br>

[Morphological rendering of brain ganglia](celltype_compendium/celltype_RGLs/brain_ganglia.html){.external target="_blank"} of the three-day-old *Platynereis dumerilii* larva.
<br>

[Morphological rendering of the mushroom bodies](celltype_compendium/celltype_RGLs/mushroom_bodies.html){.external target="_blank"} of the three-day-old *Platynereis dumerilii* larva.
<br>

[Morphological rendering of the mechanosensory girdle](celltype_compendium/celltype_RGLs/mechanosensory_girdle.html){.external target="_blank"} of the three-day-old *Platynereis dumerilii* larva.
<br>

[Cell type compendium](celltype_compendium/index.html){.external target="_blank"} of the three-day-old *Platynereis dumerilii* larva.
<br>

[Interactive visualisation of the connectome graph](Full_connectome_modules_with_names.html){.external target="_blank"} of the three-day-old *Platynereis dumerilii* larva, containing all cells in the whole-body connectome. Nodes represent individual neurons or effector cells. Directed edges (arrows) represent synaptic connections from the presynaptic neuron to the postsynaptic cell. Nodes are coloured by connectome module.
<br>

[Interactive visualisation of the cell-type-level connectome graph](Figure4_celltype_network.html){.external target="_blank"} of the three-day-old *Platynereis dumerilii* larva. Nodes represent cell types (cells grouped into one node by cell type).
<br>

[Interactive visualisation of the cell-type-level connectome graph](network_with_transmitters.html){.external target="_blank"}, same graph as before, but the nodes are coloured by transmitter phenotype.
<br>

[Interactive visualisation of the cell-type-level connectome graph](network_with_IO_layout.html){.external target="_blank"}, same graph as before, but the nodes are arranged based on the ratio of incoming to outgoing synapses.